Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses

### 2021-04-12 manuscript/manuscript.Rmd
Add thing about PTS ceiling effect

---
title: "Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "April 10, 2021"
published: This is an early draft of a working paper.
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  Many countries across the globe use legal restrictions to crack down and limit civil society. These attempts to repress civil society organizations—many of which are sanctioned by law—represent a bureaucratic form of repression that could indicate more severe human rights abuses and democratic backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we explore if crackdowns on NGOs predict broader human rights repression. We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of political terror, physical violence, and violations of private civil liberties. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deterioration in space for civil society.
keywords: "civil society restrictions, human rights, political terror, civil liberties"
reference-section-title: References
bibliography: bibliography.bib
title-page: true
endnotes: false
endfloat: false
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

As new laws are added, though, the distribution of probabilities for each of the PTS levels shifts. Adding new barriers to advocacy, entry, and funding decreases the likelihood of seeing Level 1 and increases the probability of a country turning to either Level 2 or Level 3 (with extensive political imprisonment and trial-free detention). The marginal effects of all forms of NGO laws results in this turn from Level 1 to Levels 2 or 3, but this switching effect is strongest following additional barriers to advocacy. Anti-NGO laws do not predict more dramatic shifts in the PTS. Across all counts of legal barriers, the overall probability of seeing Levels 4 and 5 is minuscule, and the adding new laws does little to shift that probability.

>>>>

---
title: "Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "April 12, 2021"
published: This is an early draft of a working paper.
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  Many countries across the globe use legal restrictions to crack down and limit civil society. These attempts to repress civil society organizations—many of which are sanctioned by law—represent a bureaucratic form of repression that could indicate more severe human rights abuses and democratic backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we explore if crackdowns on NGOs predict broader human rights repression. We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of political terror, physical violence, and violations of private civil liberties. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deterioration in space for civil society.
keywords: "civil society restrictions, human rights, political terror, civil liberties"
reference-section-title: References
bibliography: bibliography.bib
title-page: true
endnotes: false
endfloat: false
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

As new laws are added, though, the distribution of probabilities for each of the PTS levels shifts. Adding new barriers to advocacy, entry, and funding decreases the likelihood of seeing Level 1 and increases the probability of a country turning to either Level 2 or Level 3 (with extensive political imprisonment and trial-free detention). The marginal effects of all forms of NGO laws results in this turn from Level 1 to Levels 2 or 3, but this switching effect is strongest following additional barriers to advocacy. Anti-NGO laws do not predict more dramatic shifts in the PTS. Across all counts of legal barriers, the overall probability of seeing Levels 4 and 5 is minuscule, and the adding new laws does little to shift that probability. This is not unexpected however—the PTS has a natural ceiling at Level 5, and few countries that engage in such severe repression compared to lower levels of the index. The majority of movement across PTS categories should thus occur between Levels 1–3.
